---
title: "Untitled"
Author: "Sarah Castellon"
output: html_document
date: "2024-09-27"
---


This assignment has three questions:  

1. in a class of 35 students, Dr. J bets $1 against student's $1  that at least 2 students have the same birthday, month, day.
Find the probability that at least two student have the same birthday.  If 26 students in the class take the bet how much is Dr. J expected to win/lose.   How much should Dr. J wager against the student's $1 to make this a fair bet?

2. Assume class size is 26 instead of 35.  With the same bet ($1 ot $1) what is the probability Dr. J wins the bet now.  If 20 of the 26 students make the bet, what is Dr. J's expected profit/loss?

3. What size class (how many students) makes the $1 againt $1 a fair bet?

##Question 1: answer below   


```{r, echo=TRUE}
# Set up parameters
set.seed(123)  # For reproducibility
n <- 35  # Class size
simulations <- 100  # Number of simulations
same_birthday <- numeric(simulations)  # To store probabilities

# Function to check if two or more people share the same birthday
has_same_birthday <- function(class_size) {
  birthdays <- sample(1:365, class_size, replace = TRUE)
  any(duplicated(birthdays))
}

# Simulate the problem 100 times
for (i in 1:simulations) {
  # Repeat the simulation 10,000 times for each i
  count <- 0
  for (j in 1:10000) {
    if (has_same_birthday(n)) {
      count <- count + 1
    }
  }
  same_birthday[i] <- count / 10000  # Save the probability for this simulation
}

# Plot a histogram of the probabilities
hist(same_birthday, main = "Histogram of Birthday Problem Probabilities", 
     xlab = "Probability", col = "skyblue", border = "white")

# Calculate and print the mean probability
mean_prob <- mean(same_birthday)
cat("mean =", mean_prob, "\n")

# Calculate 95% confidence interval
conf_interval <- quantile(same_birthday, probs = c(0.025, 0.975))
cat("95% Confidence Interval:", conf_interval, "\n")
```

```{r}
EV= mean_prob*26-(1-mean_prob)*26
print(paste("ev=",EV))
```

```{r}
FAIR=(mean_prob)/(1-mean_prob)
print(paste("fairbet=", FAIR))
```

The probability that at least two students have the same birthday is .81. Dr. J expected to win.   Dr. J should wager 4.39 against the student's $1 to make this a fair bet.

##Question 2: answer below

```{r, echo=TRUE}
# Set up parameters
set.seed(123)  # For reproducibility
n <- 26  # Class size
simulations <- 100  # Number of simulations
same_birthday <- numeric(simulations)  # To store probabilities

# Function to check if two or more people share the same birthday
has_same_birthday <- function(class_size) {
  birthdays <- sample(1:365, class_size, replace = TRUE)
  any(duplicated(birthdays))
}

# Simulate the problem 100 times
for (i in 1:simulations) {
  # Repeat the simulation 10,000 times for each i
  count <- 0
  for (j in 1:10000) {
    if (has_same_birthday(n)) {
      count <- count + 1
    }
  }
  same_birthday[i] <- count / 10000  # Save the probability for this simulation
}

# Plot a histogram of the probabilities
hist(same_birthday, main = "Histogram of Birthday Problem Probabilities", 
     xlab = "Probability", col = "skyblue", border = "white")

# Calculate and print the mean probability
mean_prob <- mean(same_birthday)
cat("mean =", mean_prob, "\n")

# Calculate 95% confidence interval
conf_interval <- quantile(same_birthday, probs = c(0.025, 0.975))
cat("95% Confidence Interval:", conf_interval, "\n")
```

```{r}
EV= mean_prob*20-(1-mean_prob)*20
print(paste("ev=",EV))
```
With the same bet ($1 ot $1) the probability Dr. J wins the bet now is .60.  If 20 of the 26 students make the bet, Dr. J's expected profit is 3.91.


##Question3: Answer below 

```{r, echo=TRUE}
# Set up parameters
set.seed(123)  # For reproducibility
n <- 23  # Class size
simulations <- 100  # Number of simulations
same_birthday <- numeric(simulations)  # To store probabilities

# Function to check if two or more people share the same birthday
has_same_birthday <- function(class_size) {
  birthdays <- sample(1:365, class_size, replace = TRUE)
  any(duplicated(birthdays))
}

# Simulate the problem 100 times
for (i in 1:simulations) {
  # Repeat the simulation 10,000 times for each i
  count <- 0
  for (j in 1:10000) {
    if (has_same_birthday(n)) {
      count <- count + 1
    }
  }
  same_birthday[i] <- count / 10000  # Save the probability for this simulation
}

# Plot a histogram of the probabilities
hist(same_birthday, main = "Histogram of Birthday Problem Probabilities", 
     xlab = "Probability", col = "skyblue", border = "white")

# Calculate and print the mean probability
mean_prob <- mean(same_birthday)
cat("mean =", mean_prob, "\n")

# Calculate 95% confidence interval
conf_interval <- quantile(same_birthday, probs = c(0.025, 0.975))
cat("95% Confidence Interval:", conf_interval, "\n")
```

```{r}
EV= mean_prob*20-(1-mean_prob)*20
print(paste("ev=",EV))
```
The class size would have to be 23 to make the $1 againt $1 a fair bet